---
title: "PA1_Template"
author: "Hayk Grigoryan"
date: "January 14, 2017"
output: md_document
---
##Setting global options
```{r setoptions, echo = TRUE}
knitr::opts_chunk$set(echo = TRUE)
```

##Calling required libraries
```{r libraries, results = "hide"}
library(dplyr)
library(ggplot2)
library(lubridate)
library(lattice)
library(knitr)
```
##Reading data to R

```{r reading}
data <- read.csv("./activity.csv")
df <- tbl_df(data)
df$date <- as.Date(df$date)
```
--
##Calculating the summary and making a histogram

```{r summary}
mean <- mean(df$steps, na.rm = TRUE)
median <- median(df$steps, na.rm = TRUE)
s <- df %>% group_by(date) %>% summarize(daily_steps = sum(steps, na.rm = TRUE))
hist(s$daily_steps, breaks = 20, main = "Histogram of Steps", xlab = "Daily Steps")
```
The total number of steps taken each day 
- Mean is `r mean` 
- Median is `r median`

##Daily walking pattern
```{r pattern}
p <- df %>% group_by(interval) %>% summarize(steps = mean(steps, na.rm = TRUE))
p$minute <- p$interval %% 100 
p$hour <- p$interval %/% 100
p$time <- hm(paste(p$hour,p$minute, sep = ":"))
p$time <- as_date(p$time)
plot(p$time, p$steps, type = "l" )
max <- p[which(p$steps == max(p$steps)),]$time
hour <- hour(max)
minute <- minute(max)
time <- sprintf("%s:%s",hour, minute)
```

The interval of the day when the person is most active in walking is `r time`

##Imputing missing values
```{r impute_missing}
miss <- sum(is.na(df$steps))
impute.median <- function(x) replace(x, is.na(x), median(x, na.rm = TRUE))
df <- df  %>% mutate(steps = impute.median(steps))
```
Now when we calculated the missing values using the median values of each interval 
it is time to see how it changed the median and mean values of the steps 
taken each day.

```{r summary_full}
mean <- mean(df$steps, na.rm = TRUE)
median <- median(df$steps, na.rm = TRUE)
s <- df %>% group_by(date) %>% 
        summarize(daily_steps = sum(steps, na.rm = TRUE))
hist(s$daily_steps, breaks = 20, main = "Histogram of Steps", xlab = "Daily Steps")
```

The total number of steps taken each day when missing values are filled
- Mean is `r mean` 
- Median is `r median`


##Looking at weekend effect

```{r weekend}
df <- df %>% 
        mutate(weekday = ifelse(weekdays(date) 
                %in% c("Saturday", "Sunday"), "Weekend", "Weekday"))
p <- df %>% group_by(interval, weekday) %>% 
        summarize(steps = mean(steps, na.rm = TRUE))
xyplot(steps ~ interval | weekday, p, type = 'l', layout = c(1,2))
```