---
title: "PA1_Template"
author: "Hayk Grigoryan"
date: "February 02, 2017"
output: md_document
---
##Setting global options
```{r setoptions, echo = TRUE}
knitr::opts_chunk$set(echo = TRUE)
```

##Calling required libraries
```{r libraries, results = "hide"}
library(dplyr)
library(ggplot2)
library(lubridate)
library(lattice)
library(knitr)
```
##Reading data to R

```{r reading}
data <- read.csv("./activity.csv")
df <- tbl_df(data)
df$date <- as.Date(df$date)
```

##Calculating the summary and making a histogram

```{r summary}
s <- df %>% group_by(date) %>% summarize(daily_steps = sum(steps, na.rm = TRUE))
mean <- mean(s$daily_steps)
median <- median(s$daily_steps)
hist(s$daily_steps, breaks = 20, main = "Histogram of Steps", xlab = "Daily Steps")
```

The total number of steps taken each day 

* Mean      `r format(round(mean))` 
* Median     `r format(round(median))`

##Daily walking pattern
```{r pattern}
p <- df %>% group_by(interval) %>% summarize(steps = mean(steps, na.rm = TRUE))
p$minute <- p$interval %% 100 
p$hour <- p$interval %/% 100
p$time <- hm(paste(p$hour,p$minute, sep = ":"))
p$time <- as_date(p$time)
plot(p$time, p$steps, type = "l" )
max <- p[which(p$steps == max(p$steps)),]$time
hour <- hour(max)
minute <- minute(max)
time <- sprintf("%s:%s",hour, minute)
```

The interval of the day when the person is most active in walking is morning time `r time`

##Imputing missing values

We input missing values using the mean values of each interval 
```{r impute_missing}
miss <- sum(is.na(df$steps))
impute.mean <- function(x) replace(x, is.na(x), mean(x, na.rm = TRUE))
dk <- df  %>% mutate(steps = impute.mean(steps))
```
Now when we calculated the missing values it is time to see how it changed the median and mean values of the steps taken each day.

```{r summary_full}
sk <- dk %>% group_by(date) %>% 
        summarize(daily_steps = sum(steps, na.rm = TRUE))
mean_f <- mean(sk$daily_steps)
median_f <- median(sk$daily_steps)
hist(s$daily_steps, breaks = 20, main = "Histogram of Steps", xlab = "Daily Steps")
```

The total number of steps taken each day when missing values are filled

* Mean      `r format(round(mean_f))` 
* Median    `r format(round(median_f))`


##Looking at weekend effect

At weekend people wake up later and their walking activity is much dispersed throughout a day

```{r weekend}
df <- df %>% 
        mutate(weekday = ifelse(weekdays(date) 
                %in% c("Saturday", "Sunday"), "Weekend", "Weekday"))
p <- df %>% group_by(interval, weekday) %>% 
        summarize(steps = mean(steps, na.rm = TRUE))
xyplot(steps ~ interval | weekday, p, type = 'l', layout = c(1,2))
```